---
title: "Models - Jeffrey"
author: "Jeffrey Hsu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data & Prepare Data Structure


```{r message=FALSE, warning=FALSE}
library(car); require(dplyr); library(Hmisc) 
library(mcprofile); library(ggplot2); library(gridExtra) 
library(reshape2); library(GGally); library(ggcorrplot)
library(package = MASS) # for ordinal response variable modeling

dehart <- read.table(file="DeHartSimplified.csv", header=TRUE, sep=",")
dehart$dayweek_f <- factor(dehart$dayweek); levels(dehart$dayweek_f) = c("mon","tue","wed","thu","fri","sat","sun")
dehart$gender_f <- factor(dehart$gender); levels(dehart$gender_f) = c("male","female")
# nrel_boolean
#hist(dehart$nrel, breaks=0:10-0.5)
dehart_nrel = dehart[which(dehart$nrel != 0),]
dehart$nrel_bool = 1*(dehart$nrel > 0)
dehart$rosn_cat <- cut(dehart$rosn, breaks=c(-1, 2.8, 3.4, Inf), labels = c("low", "mid", "high"))
dehart$trel <- dehart$nrel/sd(dehart$nrel) + dehart$prel/sd(dehart$prel)

#hist(dehart$rosn, breaks=seq(2,4.3,0.1)-0.1)
#hist(dehart$posevent)
#hist(dehart$desired, breaks=seq(0,9,0.1))


dehart[is.na(dehart$numall) | is.na(dehart$desired), ] # id 2, 42, 110, 116, dayweek 7, 7, 1, 3
dehart_cleaned <- dehart[!is.na(dehart$numall) & !is.na(dehart$desired),]
```

There're 2 outcome variable we are interested: number of drinks (\textbf{numall}) and the desire to drink (\textbf{desired}). Given \textbf{numall} is a count variable, the natural model is poisson regression. The \textbf{desired} variable is in range 1-8 with 22 distinct values. The distribution is rather smooth and steady, however, given there're only 22 distinct variables and the fact that the score is bounded, we cannot simply treat this as a continuous variable and use OLS for it. We decided to model this variable with ordinal response regression model. Each integar value range (e.g. 1-2, 2-3... etc.) is used as a level. 

These 2 outcome variables are modeled with the following different set of explanatory variables:

## Base Model

In the base model we use the boolean version of negative relationship variable (\textbf{nrel_bool}), trait self-esteem (\textbf{rosn}) and their interaction as explanatory variables. \textbf{nrel_bool} is a binary variable with 1 indicating presence of negative relationship and 0 elsewise.

```{r}
## numall poisson regression
base_numall <- glm(numall ~ nrel_bool + rosn + nrel_bool*rosn, 
                   family = poisson(link = "log"), data = dehart_cleaned)
summary(base_numall)

base_numall.ha <- glm(numall ~ nrel_bool + rosn, 
                   family = poisson(link = "log"), data = dehart_cleaned)

Anova(base_numall)
anova(base_numall.ha, base_numall)

## desired ordinal regression
dehart_cleaned$desired_ord <- factor(cut(dehart_cleaned$desired, 
                                         breaks=c(0.0009, 2.0001, 3.0001, 4.0001, 5.0001, 6.0001, 7.0001, 8.0001),
                                         labels=c("1-2", "2-3", "3-4", "4-5", "5-6", "6-7", "7-8")))
levels(dehart_cleaned$desired_ord)
base_desired <- polr(formula = desired_ord ~ nrel_bool + rosn + nrel_bool*rosn, 
                     data = dehart_cleaned, method ="logistic")
summary(base_desired)
Anova(base_desired)
```

## Intermediate Model

```{r}
## numall poisson regression
inter_numall <- glm(numall ~ nrel_bool + rosn + nrel_bool*rosn + prel + prel*rosn + dayweek_f,
                    family = poisson(link = "log"), data = dehart_cleaned)
summary(inter_numall)
Anova(inter_numall)

## desired ordinal regression
inter_desired <- polr(formula = desired_ord ~ nrel_bool + rosn + nrel_bool*rosn + prel + prel*rosn + dayweek_f, 
                      data = dehart_cleaned, method ="logistic")
summary(inter_desired)
Anova(inter_desired)
```


## Full Model

```{r}
## numall poisson regression
full_numall <- glm(numall ~ nrel_bool + rosn + nrel_bool*rosn + prel + prel*rosn +
                     dayweek_f + age + gender + negevent + posevent + state,
                   family = poisson(link = "log"), data = dehart_cleaned)
summary(full_numall)

full_numall.ha <- glm(numall ~ nrel_bool + rosn + prel +
                     dayweek_f + age + gender + negevent + posevent + state,
                   family = poisson(link = "log"), data = dehart_cleaned)
anova(full_numall, full_numall.ha)

## desired ordinal regression
full_desired <- polr(formula = desired_ord ~ nrel_bool + rosn + nrel_bool*rosn + prel + prel*rosn +
                       dayweek_f + age + gender + negevent + posevent + state, 
                      data = dehart_cleaned, method ="logistic")
summary(full_desired)

full_desired.ha <- polr(formula = desired_ord ~ nrel_bool + rosn + prel +
                       dayweek_f + age + gender + negevent + posevent + state, 
                      data = dehart_cleaned, method ="logistic")
anova(full_desired, full_desired.ha)
```